---
title: "Lab 5"
author: "Kelsey Leach"
date: "`r format(Sys.Date())`"
output: github_document
---

Lab Report 5: The purpose of this lab is to learn to use the rename, mutate, summarize and grouping functions and to learn to make a scatterplot and a strip plot with grouped data.

1. Rename each variable so that it is all lower-case and uses an underscore _ instead of a period . in the name (the recommended coding style in the tidyverse style guide). Print the resulting table.

```{r}
library(tidyverse)
iris <- as_tibble(iris)
iris <- rename(iris, sepal_length = Sepal.Length, sepal_width = Sepal.Width, petal_length = Petal.Length, petal_width = Petal.Width, species = Species)
iris
```

2. Convert the four numerical variables from cm to mm by multiplying by 10. Print the resulting table.
```{r}
iris2 <- mutate(iris, sepal_length_cm = sepal_length * 10, sepal_width_cm = sepal_width * 10, petal_length_cm = petal_length * 10, petal_width_cm = petal_width * 10)
iris2 <- select(iris2, species, sepal_length_cm, sepal_width_cm, petal_length_cm, petal_width_cm)
iris2
```

3. Calculate sepal area and petal area (area is equal to length multiplied by width). Print a table with only the variables sepal area, petal area, and species.
```{r}
iris3 <- mutate(iris, sepal_area = sepal_length * sepal_width, petal_area = petal_length * petal_width)
iris3 <- select(iris3, sepal_area, petal_area, species)
iris3
```

4. Calculate the following statistics for the entire dataset from the sepal length variable and print the resulting table: sample size, maximum value, minimum value, range, median, first quartile (q1), third quartile (q2), and inter-quartile range (iqr).
```{r}
summary4 <- summarise(iris, sampl_size = n(), max_value = max(sepal_length), min_value = min(sepal_length), range = max_value - min_value, median = median(sepal_length), q1 = quantile(sepal_length, probs = 0.25), q3 = quantile(sepal_length, probs = 0.75), iqr = q3 - q1)
summary4
```

5. Calculate the following statistics for each species from the petal width variable and print the resulting table: sample size, mean, standard deviation, variance, standard error of the mean, and approximate 95% confidence interval
```{r}
iris_grouped <- group_by(iris, species)
summary5 <- summarise(iris_grouped, sampl_size = n(), mean_petal_width = mean(petal_width), sd_petal_width = sd(petal_width), variance = var(petal_width), sem = mean_petal_width / sqrt(sampl_size), ci_lower = mean_petal_width - 2 * sem, ci_upper = mean_petal_width + 2 * sem)
summary5
```

6. Visualize the relationship between petal length and species using a strip plot.
```{r}
ggplot(data = iris) +
  geom_jitter(mapping = aes(x = species, y = petal_length))  +
  labs(x = "Species", y = "Petal Length (mm)")
```

7. Starting with the previous graph, add the mean and 95% confidence interval for each species
```{r}
iris_summary <-
  summarize(
    iris_grouped, 
    mean_petal_length = mean(petal_length),
    sem = sd(petal_length) / sqrt(n()),
    upper_limit = mean_petal_length + 1.96 * sem,
    lower_limit = mean_petal_length - 1.96 * sem
  )
ggplot(data = iris) +
  geom_jitter(mapping = aes(x = species, y = petal_length)) +
  geom_crossbar(
    data = iris_summary, 
    mapping = aes(x = species, y = mean_petal_length, ymax = upper_limit, ymin = lower_limit),
    color = "red") +
    labs(x = "Species", y = "Petal Length (mm)")
```

8. Visualize the relationship between petal length, petal width, and species using a scatterplot. Map the two numerical variables to the x and y axes and map species to the color and shape aesthetics.
```{r}
ggplot(data = iris) +
  geom_point(mapping = aes(x = petal_length, y = petal_width, color = species), alpha = 0.8) + labs(x = "Petal Length (mm)", y = "Petal Width (mm)", color = "Species") 

```

Here is my session information:

```{r}
sessioninfo::session_info()
```